An Accurate Model for SARS-CoV-2 pooled RT-PCR Tests Errors

### 2021-03-21 comment.tex
wordz
intro

\section{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in breaking transmission chains and mitigating the
COVID-19 pandemic. As such, the need for large-scale testing has
resulted in the development of techniques to increase the the
throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques are
called \emph{pooling} and originated in the seminal work of Dorfman
pooling \cite{DorfmanOriginal, DorfmanYuvalDor}: Select $N$
individuals and perform a single RT-PCR test on their combined
(\emph{pooled}) samples. If the pooled test yields a positive result
--- test each individual separately. The throughput efficiency of
Dorfman pooling for has been demonstrated empirically
\cite{DorfmanYuvalDor} and its error rates thoroguhly investigated
\cite{Kim, Simplistic1, OptimalDorfmanPool}.

Pooling RT-PCR tests is not perfect, and there is always some nonzero
probability of error. Typically, authors assume the probability of a
true-positive (the test's \emph{sensitivity}) does not depend on the
number of infected samples in the pool, but rather on the existence of
at least one such sample. Thus, the probability of a positive result
in a pooled test is (assumed) identical for a pool with one sample
from an infected individual and (e.g.) five such samples. This
assumption is common in the group testing literature \cite{Kim,
  OptimalDorfmanPool}, as well as in more specific, COVID-19 focused
studies \cite{Simplistic1, Simplistic2}. In this study we refute these
commonly made assumptions and show how they affect estimation of
false-negative rates for Dorfman pooling.

@@

To summarize, we assume that for a single pool, a positive test result
is generated in one of two paths. Either, SARS-CoV-2 RNA from an
infected individual's sample is correctly amplified and detected, and
this can happen for each positive sample in a pool, with probability
$\Se$. Or, some erroneous amplification occurs (e.g. contaminant viral
RNA is introduced), an event that occurs at most once per pool, with
probability $1-\Sp$. Our model is illustrated in
Figure~\ref{fig:likelihood}, and summarized in
%% A negative pooled test result occurs when all detection paths (both
%% correct and erroneous) fail. The probability of no false-detection
%% accounts for the $\spec$ term. The probability of no correct detection
%% is $(1-\sens)$ per infected individual. The probability that all such
%% paths fail is the product of the above mentioned terms, displayed in
\eqref{eq:likelihood a}.%%  The probability of a positive result,
%% presented in \eqref{eq:likelihood b}, is simply the complement.
%% The probability of a negative pooled test result is presented in
%% \eqref{eq:likelihood a} (along with its complement
%% \eqref{eq:likelihood b}), and explained below.
%\begin{subequations}
\begin{align}%\label{eq:single likelihood}
  %\begin{split}
    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum\theta_i}. \label{eq:likelihood a}
    %
    %
    %
    %% \Pr(\data=1 | \theta) &= 1 - \spec\prod_{i=1}^N
    %% (1-\sens)^{\theta_i} = 1 - \spec(1-\Se)^{\sum\theta_i} . \label{eq:likelihood b}
  %\end{split}
\end{align}
%\end{subequations}
%% Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
%% recalling that $\data \in \{0,1\}$ yields:

@@

\subsection{Application}
As an application of our model, we calculate the differnece in
false-negative rate for a single \emph{infected} individual,
henceforth referred to as "Donald".  We distinguish three types of
false-negative events when performing pooling. A \emph{single test's}
false-negative is the event of a negative result upon testing Donald
separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
false-negative occurs when a pooled test containing Donald's sample
(and other samples) yields a negative result, i.e., the pooling fails
to detect at least one positive result. Lastly, a \emph{scheme}
false-negative occurs when an entire pooling scheme fails to identify
Donald as infected. Our goal is to calculate Dorfman's scheme
false-negative rate. Rephrasing, we wish to answer the following
question: what is the probability of not identifying Donald as
infected under a Dorfman pooling scheme?

>>>>

\section{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in breaking transmission chains and mitigating the
COVID-19 pandemic. As such, the need for large-scale testing has
resulted in the development of techniques to increase the the
throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques, often termed
 \emph{pooling}, or group testing,  have originated in the seminal work of Dorfman et al. and their pooling technique
 \cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one selects $N$
individuals and performs a single RT-PCR test on their combined
(\emph{pooled}) samples. If the pooled test yields a positive result, then  each individual is retested separately; otherwise, everyone is declared negative. The throughput efficiency of
Dorfman pooling for has been demonstrated empirically
\cite{DorfmanYuvalDor} and its error rates thoroguhly investigated
\cite{Kim, Simplistic1, OptimalDorfmanPool}.

Pooling RT-PCR tests is not perfect, and there is always some nonzero
probability of error. Typically, previous research assumed that the probability of a true-positive (the test's \emph{sensitivity}) does not depend on the
number of infected samples in the pool, but rather on the existence of
at least one such sample. Thus, the probability of a positive result
in a pooled test has been (assumed) identical for a pool with one sample
from an infected individual and, e.g., five such samples. This
assumption is common in the group testing literature \cite{Kim,
  OptimalDorfmanPool}, as well as in more specific, COVID-19 focused
studies \cite{Simplistic1, Simplistic2}. In this study we refute these
commonly made assumptions and show how using a more accurate probabilistic model affects estimation of
false-negative rates for Dorfman pooling.

@@

To summarize, we assume that for a single pool, a positive test result
is generated in one of two paths. Either, SARS-CoV-2 RNA from an
infected individual's sample is correctly amplified and detected, and
this can happen for each positive sample in a pool, with probability
$\Se$. Or, some erroneous amplification occurs (e.g. contaminant viral
RNA is introduced), an event that occurs at most once per pool, with
probability $1-\Sp$. Our model is illustrated in
Figure~\ref{fig:likelihood}, and summarized in
%% A negative pooled test result occurs when all detection paths (both
%% correct and erroneous) fail. The probability of no false-detection
%% accounts for the $\spec$ term. The probability of no correct detection
%% is $(1-\sens)$ per infected individual. The probability that all such
%% paths fail is the product of the above mentioned terms, displayed in
\eqref{eq:likelihood a}.%%  The probability of a positive result,
%% presented in \eqref{eq:likelihood b}, is simply the complement.
%% The probability of a negative pooled test result is presented in
%% \eqref{eq:likelihood a} (along with its complement
%% \eqref{eq:likelihood b}), and explained below.
%\begin{subequations}
\begin{align}%\label{eq:single likelihood}
  %\begin{split}
    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
    %
    %
    %
    %% \Pr(\data=1 | \theta) &= 1 - \spec\prod_{i=1}^N
    %% (1-\sens)^{\theta_i} = 1 - \spec(1-\Se)^{\sum\theta_i} . \label{eq:likelihood b}
  %\end{split}
\end{align}
%\end{subequations}
%% Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
%% recalling that $\data \in \{0,1\}$ yields:

@@

\subsection{Application}
As an application of our model, we calculate the differnece in
false-negative rate for a single \emph{infected} individual,
henceforth referred to as "Donald".  We distinguish three types of
false-negative events when performing pooling. A \emph{single test}'s
false-negative is the event of a negative result upon testing Donald
separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
false-negative occurs when a pooled test containing Donald's sample
(and other samples) yields a negative result, i.e., the pooling fails
to detect at least one positive result. Lastly, a \emph{scheme}
false-negative occurs when an entire pooling scheme fails to identify
Donald as infected. Our goal is to calculate Dorfman's scheme
false-negative rate. Rephrasing, we wish to answer the following
question: what is the probability of not identifying Donald as
infected under a Dorfman pooling scheme?
